---
title: "Descriptive stats plots"
author: "Molly Martorella"
date: "12/3/2019"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)

theme_set(theme_bw())
```

# Data

## NYC 311

Load and tidy the data. Data is 100k randomly sampled complaints from all neighborhoods across all NYC boroughs (600k total samples). Years sampled were 2014-2018.

Data downloaded from:
https://wetransfer.com/downloads/f8c5d6c17483e279ff56018db9c44cc420191201000026/c5f48b

```{r, warning=FALSE}

nyc <- read_csv(file = "./p8105nyc_311_100k.csv") %>% 
    janitor::clean_names()

nyc_tidy <- nyc %>% 
    filter(borough != "Unspecified") %>% 
    separate(closed_date, 
             into = c("closed_month","closed_day","closed_year"), 
             sep = "\\/" ) %>%
    separate(closed_year, 
             into = c("closed_year","closed_time"), 
             sep = " ") %>% 
    mutate(
        created_year = as.numeric(created_year),
        created_month = as.numeric(created_month),
        created_day = as.numeric(created_day),
        city = as.factor(city),
        status = as.factor(status),
        borough = as.factor(borough),
        agency = as.factor(agency),
        complaint_type = as.factor(complaint_type),
        community_board = as.factor(community_board),
       open_complaint = ifelse(status == "Closed", yes = 0, no = 1),
      # open_complaint = ifelse(is.na(closed_year),  yes = 1, no = 0),
        complaint_simp = case_when(
            str_detect(complaint_type, 
                       regex("street", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("sidewalk", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("curb", ignore_case = TRUE)) ~ "Street Condition",
            str_detect(complaint_type, 
                       regex("noise", ignore_case = TRUE)) ~ "Noise",
            str_detect(complaint_type, 
                       regex("heat", ignore_case = TRUE)) ~ "Heat",
            str_detect(complaint_type, 
                       regex("water", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("leak", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("plumbing", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("boiler", ignore_case = TRUE)) ~ "Water/plumbing",
            str_detect(complaint_type, 
                       regex("paint", ignore_case = TRUE)) ~ "Paint/Plaster",
            str_detect(complaint_type, 
                       regex("asbestos", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("lead", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("hazard", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("mold", ignore_case = TRUE)) ~ "Hazard Material",
            str_detect(complaint_type, 
                       regex("elevator", ignore_case = TRUE)) 
            |str_detect(complaint_type, 
                        regex("maintenance", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("electric", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("stairs", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("window", ignore_case = TRUE)) 
            |str_detect(complaint_type, 
                        regex("appliance", ignore_case = TRUE)) ~ "Maintenance",
            str_detect(complaint_type, 
                       regex("sanita", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("rodent", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("dirty", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("sew", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("standing water", ignore_case = TRUE)) ~ "Sanitation",
            str_detect(complaint_type, 
                       regex("tree", ignore_case = TRUE)) ~ "Tree",
            str_detect(complaint_type, 
                       regex("parking", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("car", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("drive", ignore_case = TRUE))
            |str_detect(complaint_type,  
                        regex("vehicle", ignore_case = TRUE))
            |str_detect(complaint_type,  
                        regex("traffic", ignore_case = TRUE)) ~ "Car/Traffic",
            str_detect(complaint_type, 
                       regex("air", ignore_case = TRUE)) ~ "Air Quality",
            str_detect(complaint_type, 
                       regex("collection", ignore_case = TRUE)) ~ "Collection",
            str_detect(complaint_type, 
                       regex("homeless", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("panhandling", ignore_case = TRUE)) ~ "Homeless"),
        health_complaint = ifelse(
            complaint_simp %in% c("Heat", "Water/Plumbing", "Hazard Material", "Sanitation", "Air Quality"), yes = 1, no = 0),
        complaint_simp = as.factor(complaint_simp),
        open_health_complaint = case_when(
            open_complaint == 1 & health_complaint == 1 ~ 1,
            open_complaint == 0 | health_complaint == 0 ~ 0
        ),
       # openCorr = ifelse(status == "Closed", yes = 0, no = 1),
        status = as.factor(status)
    )

```

Used skimr::skim, colnames, and levels(factor(nyc$complaint_type)) to investigate and tidy the data. New variables were created to establish a binary status for closed (0) or open (1) complaints.

Key variables include: `year`, `borough`, `community_board`, `complaint_type`, and `status`.

Newly added variables include:

1. `complaint_simp` - based on key words, condenses complaint types into the following categories: Air Quality, Car/Traffic, Collection, Hazard Material, Heat, Homeless, Maintenance, Noise, Paint/Plaster, Sanitation, Street Condition, Tree, and Water/plumbing.

2. `health_complaint` - binary yes (1) or no (0), based on health associated categories within `complaint_simp`.

3. `open_health_complaint` - binary categorization, either health complaint that is open (1), or non-health related complaint or closed health complaint (0).

4. `open_complaint` - binary categorization, closed `status` (0), open (1)

## Income and population characteristics data:

The variable, `community_board`, will be used to left join NYC 311 data with income and population characteristics data sourced from American Community Survey by Community District.

```{r}

inc_df = read_csv("./Med_income_2017.csv") %>% 
    janitor::clean_names() %>% 
    mutate(
        inc_1000s = round(median_income/1000, 1),
        income_bracket = case_when(
            median_income >= 20000 & median_income <= 30000 ~ "20k-30k",
            median_income > 30000 & median_income <= 40000 ~ "30-40k",
            median_income > 40000 & median_income <= 50000 ~ "40-50k",
            median_income > 50000 & median_income <= 60000 ~ "50-60k",
            median_income > 60000 & median_income <= 70000 ~ "60-70k",
            median_income > 70000 & median_income <= 80000 ~ "70-80k",
            median_income > 80000 & median_income <= 90000 ~ "80-90k",
            median_income > 90000 & median_income <= 100000 ~ "90-100k",
            median_income > 100000 & median_income <= 125000 ~ "100-125k",
            median_income > 125000 & median_income <= 150000 ~ "125k+",
        ),
        income_bracket = as.factor(income_bracket),
        income_bracket = fct_relevel(income_bracket, c("20k-30k", "30-40k", "40-50k", "50-60k", "60-70k", "70-80k", "80-90k", "90-100k", "100-125k", "125k+"))
    )


nyc_inc = left_join(nyc_tidy, inc_df, by = "community_board")

```

`income_bracket` and `inc_1000s` (rounds income and divides by 1000) variables were added to the neighborhood population characteristics data to ease downstream plots and analyses. The data were left joined using the `community_board` variable.

## Combined and tidied data for plotting

```{r}

nyc_plots <- nyc_inc %>% 
    group_by(area_name, created_year) %>%
    add_count(area_name, name = "number_complaints") %>% 
    mutate(
        num_unsolved = sum(open_complaint),
        num_health_complaint = sum(health_complaint),
        num_open_health = sum(open_health_complaint)
    ) %>% 
    select(-unique_key, -city, -park_borough, -agency, -agency_name, -descriptor, -incident_zip, -incident_address, -street_name, -cross_street_1, -cross_street_2, -intersection_street_1, -intersection_street_2, -landmark, -facility_type, -resolution_description, -resolution_action_updated_date, -bbl, -x_coordinate_state_plane, -y_coordinate_state_plane, -open_data_channel_type, -park_facility_name, -vehicle_type, -taxi_company_borough, -taxi_pick_up_location, -bridge_highway_name, -bridge_highway_direction, -bridge_highway_segment, -latitude, -longitude, -location, -road_ramp, -location_type, -address_type, -map_id)

```

The newly combined dataframe was grouped according to `area_name`, a key variable from the population characteristics dataset that provides the local neighborhood name within the borough. Extraneous variables were removed to reduce the dataframe size and ease manipulation of the data.

Newly added variables include:

1. `number_complaints` - total number of complaints within the given neighborhood the complaint was filed.

2. `num_unsolved` - total number of open/unresolved complaints within the neighborhood the complaint was filed.

3. `num_health_complaint` - total number of health related complaints within the neighborhood the complaint was filed.

4. `num_open_health` - total number of open/unresolved health complaints within the neighborhood the complaint was filed.


# Plots

#### Total number of complaints by income bracket

```{r}

# nyc_plots %>% 
#     filter(!is.na(income_bracket)) %>% 
#     group_by(income_bracket) %>% 
#     summarize(n = n(),
#               proportion = n/nrow(nyc_plots)) %>% 
#     ggplot(aes(x = income_bracket, y = proportion)) +
#     geom_col() +
#     ylab("Proportion of Total Complaints") +
#     ggtitle("Relative Number of Complaints Filed from Each Income Bracket")

nyc_plots %>% 
    filter(!is.na(income_bracket)) %>% 
    group_by(income_bracket, borough) %>% 
    summarize(n = n(),
              proportion = n/nrow(nyc_plots)) %>% 
    ggplot(aes(x = income_bracket, y = proportion, fill = income_bracket)) +
    geom_col() +
    ylab("Proportion") +
    xlab("Income Bracket") +
    ggtitle("Proportion of Total Complaints from Each Income Bracket") +
    facet_wrap(~borough, scales = "free_x") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Middle class income groups proportionally file the most complaints across all income brackets. The only exception is the Bronx, in which the proportion of 311 complaints is higher in the lowest income group.

#### Proportion of unsolved cases by income bracket

Proportion open:

```{r}

p <- nyc_plots %>% 
    filter(!is.na(income_bracket)) %>% 
    group_by(income_bracket, borough, open_complaint) %>% 
    summarize(n = n()) %>% 
    pivot_wider(names_from = open_complaint, values_from = n) %>% 
    rename(closed = '0',
           open = '1')

p %>% mutate(proportion = (open/closed)) %>% 
    ggplot(aes(x = income_bracket, y = proportion, fill = income_bracket)) +
    geom_col() +
    facet_wrap(~borough, scales = "free_x") +
    ggtitle("Proportion of Open Cases Within Each Income Bracket") +
    xlab("Income Bracket") +
    ylab("Proportion Open") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))

```


#### Total Health related complaints broken down by income bracket

```{r}
nyc_plots %>% 
    filter(!is.na(income_bracket)) %>% 
    group_by(income_bracket, borough, health_complaint) %>% 
    summarize(n = n(),
              proportion = n/nrow(nyc_plots)) %>% 
    ggplot(aes(x = income_bracket, y = proportion, fill = income_bracket)) +
    geom_col() +
    ylab("Proportion") +
    xlab("Income Bracket") +
    ggtitle("Proportion of Total Complaints from Each Income Bracket") +
    facet_wrap(~borough, scales = "free_x") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

#### Proportion Health related complaints open broken down by income bracket:

```{r}
h <- nyc_plots %>% 
    filter(!is.na(income_bracket),
           health_complaint == 1) %>% 
    group_by(income_bracket, borough, open_complaint) %>% 
    summarize(n = n()) %>% 
    pivot_wider(names_from = open_complaint, values_from = n) %>% 
    rename(closed = '0',
           open = '1')

h %>% mutate(proportion = (open/closed)) %>% 
    ggplot(aes(x = income_bracket, y = proportion, fill = income_bracket)) +
    geom_col() +
    facet_wrap(~borough, scales = "free_x") +
    ggtitle("Proportion of Health Related Open Cases Within Each Income Bracket") +
    xlab("Income Bracket") +
    ylab("Proportion Open") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

#### Median income of each complaint category:

```{r}

nyc_plots %>% 
    filter(!is.na(median_income),
           !is.na(complaint_simp)) %>% 
    mutate(complaint_simp = fct_reorder(complaint_simp, median_income)) %>% 
    ggplot(aes(x = complaint_simp, 
               y = (median_income/1000), 
               fill = complaint_simp)) +
    geom_boxplot() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          legend.position = "none") +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    ylab("Median Income (thousands)") +
    xlab("Complaint Category") +
    ggtitle("Median Income of Community District where Complaint Category Occurs")


# open cases:

nyc_plots %>% 
    filter(!is.na(median_income),
           !is.na(complaint_simp),
           open_complaint == 1) %>% 
    ggplot(aes(x = complaint_simp, 
               y = (median_income/1000), 
               fill = complaint_simp)) +
    geom_boxplot() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          legend.position = "none") +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    ylab("Median Income (thousands)") +
    xlab("Complaint Category") +
    ggtitle("Median Income of Community District where Open Complaint Category Occurs")

```


#### Not to be included:

Total complaints within each neighborhood of each borough:

```{r}

nyc_plots %>% 
    filter(!is.na(area_name)) %>% 
    group_by(area_name, borough) %>% 
    summarize(n = n()) %>% 
    ggplot(aes(x = area_name, y = n)) +
    geom_col() +
    ylab("Total Complaints") +
    xlab("Neighborhood") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
    ggtitle("Complaints by Neighborhood") +
    facet_wrap(~borough, scales = "free_x")

nyc_plots %>% 
    filter(!is.na(community_board)) %>% 
    group_by(community_board, borough) %>% 
    summarize(n = n()) %>% 
    filter(n > 100) %>% 
    ggplot(aes(x = community_board, y = n)) +
    geom_col() +
    ylab("Total Complaints") +
    xlab("Community Board") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
    ggtitle("Complaints by Community Board") +
    facet_wrap(~borough, scales = "free_x")

```

other med income plot:

```{r}
# 
# nyc_plots %>% 
#     filter(!is.na(income_bracket),
#            !is.na(complaint_simp)) %>% 
#     group_by(income_bracket, borough, complaint_simp) %>% 
#     summarize(n = n()) %>% 
#     mutate(proportion = n/nrow(nyc_plots)) %>% 
#     arrange(income_bracket, proportion) %>% 
#     mutate(order = row_number()) %>% 
#     group_by(complaint_simp) %>% 
#     ggplot(aes(x = complaint_simp, y = proportion, fill = complaint_simp)) +
#     geom_col() +
#     facet_wrap(~income_bracket, scales = "free_x", ncol = 5) +
#     theme(axis.text.x = element_text(angle = 75, hjust = 1))
# 

# 
#            complaint_simp = fct_reorder(complaint_simp, 
#                                         proportion, 
#                                         .desc = TRUE)) %>% 
    
```




# Final Plots:

## Proportion of unsolved cases by income bracket

```{r}

p <- nyc_plots %>% 
    filter(!is.na(income_bracket)) %>% 
    group_by(income_bracket, borough, open_complaint) %>% 
    summarize(n = n()) %>% 
    pivot_wider(names_from = open_complaint, values_from = n) %>% 
    rename(closed = '0',
           open = '1')

p %>% mutate(proportion = (open/closed)) %>% 
    ggplot(aes(x = income_bracket, y = proportion, fill = income_bracket)) +
    geom_col() +
    facet_wrap(~borough, scales = "free_x") +
    ggtitle("Proportion of Open Cases Within Each Income Bracket") +
    xlab("Income Bracket") +
    ylab("Proportion Open") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

## Median income of each complaint category:

```{r}

nyc_plots %>% 
    filter(!is.na(median_income),
           !is.na(complaint_simp)) %>% 
    mutate(complaint_simp = fct_reorder(complaint_simp, median_income)) %>% 
    ggplot(aes(x = complaint_simp, 
               y = (median_income/1000), 
               fill = complaint_simp)) +
    geom_boxplot() +
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          legend.position = "none") +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    ylab("Median Income (thousands)") +
    xlab("Complaint Category") +
    ggtitle("Median Income of Community District where Complaint Category Occurs")
```


## plot of percent race plotted against proportion open:

```{r}
r <- nyc_plots %>% 
    pivot_longer(cols = per_black_nh:per_other, names_to = "race", values_to = "percent_pop") %>% 
    filter(!is.na(race),
           !is.na(percent_pop)) %>% 
    group_by(race, borough, percent_pop, open_complaint) %>% 
    summarize(n = n()) %>% 
    pivot_wider(names_from = open_complaint, values_from = n) %>% 
    rename(closed = '0',
           open = '1')

r %>% mutate(proportion = (open/(open + closed))) %>% 
    ggplot(aes(x = percent_pop, y = proportion, color = race)) +
    geom_point() +
    geom_smooth(se = FALSE) +
    xlab("Percentage of Community District Population") +
    ylab("Proportion Open") +
    ggtitle("Unresolved Complaints as a Function of Neighborhood Population")
    
```

density plot of number of days to close by income group or pop proportions?










