---
title: "<span style='font-size: 25px'>Descriptive Tables</style>"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arsenal)
```


```{r message = FALSE, warning=FALSE}

#data previously combined in the nyc_open_data.Rmd file
nyc <- read_csv(file = "../p8105nyc_311_100k.csv") %>% 
    janitor::clean_names()

#creating variables that aggragate similiar complaints. 
nyc_tidy <- nyc %>%   
    filter(borough != "Unspecified") %>%  
    separate(closed_date, 
             into = c("closed_month","closed_day","closed_year"), 
             sep = "\\/" ) %>%
    separate(closed_year, 
             into = c("closed_year","closed_time"), 
             sep = " ") %>% 
    mutate(
        created_year = as.numeric(created_year),
        created_month = as.numeric(created_month),
        created_day = as.numeric(created_day),
        city = as.factor(city),
        status = as.factor(status),
        borough = as.factor(borough),
        agency = as.factor(agency),
        complaint_type = as.factor(complaint_type),
        community_board = as.factor(community_board),
        open_complaint = ifelse(status == "Closed", yes = 0, no = 1),
        # open_complaint = ifelse(is.na(closed_year),  yes = 1, no = 0),
        complaint_simp = case_when(
            str_detect(complaint_type, 
                       regex("street", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("sidewalk", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("curb", ignore_case = TRUE)) ~ "Street Condition",
            str_detect(complaint_type, 
                       regex("noise", ignore_case = TRUE)) ~ "Noise",
            str_detect(complaint_type, 
                       regex("heat", ignore_case = TRUE)) ~ "Heat",
            str_detect(complaint_type, 
                       regex("water", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("leak", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("plumbing", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("boiler", ignore_case = TRUE)) ~ "Water/plumbing",
            str_detect(complaint_type, 
                       regex("paint", ignore_case = TRUE)) ~ "Paint/Plaster",
            str_detect(complaint_type, 
                       regex("asbestos", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("lead", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("hazard", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("mold", ignore_case = TRUE)) ~ "Hazard Material",
            str_detect(complaint_type, 
                       regex("elevator", ignore_case = TRUE)) 
            |str_detect(complaint_type, 
                        regex("maintenance", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("electric", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("stairs", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("window", ignore_case = TRUE)) 
            |str_detect(complaint_type, 
                        regex("appliance", ignore_case = TRUE)) ~ "Maintenance",
            str_detect(complaint_type, 
                       regex("sanita", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("rodent", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("dirty", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("sew", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("standing water", ignore_case = TRUE)) ~ "Sanitation",
            str_detect(complaint_type, 
                       regex("tree", ignore_case = TRUE)) ~ "Tree",
            str_detect(complaint_type, 
                       regex("parking", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("car", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("drive", ignore_case = TRUE))
            |str_detect(complaint_type,  
                        regex("vehicle", ignore_case = TRUE))
            |str_detect(complaint_type,  
                        regex("traffic", ignore_case = TRUE)) ~ "Car/Traffic",
            str_detect(complaint_type, 
                       regex("air", ignore_case = TRUE)) ~ "Air Quality",
            str_detect(complaint_type, 
                       regex("collection", ignore_case = TRUE)) ~ "Collection",
            str_detect(complaint_type, 
                       regex("homeless", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("panhandling", ignore_case = TRUE)) ~ "Homeless"),
        health_complaint = ifelse(
            complaint_simp %in% c("Heat", "Water/Plumbing", "Hazard Material", "Sanitation", "Air Quality"), yes = 1, no = 0),
        complaint_simp = as.factor(complaint_simp),
        open_health_complaint = case_when(
            open_complaint == 1 & health_complaint == 1 ~ 1,
            open_complaint == 0 | health_complaint == 0 ~ 0
        ),
        # openCorr = ifelse(status == "Closed", yes = 0, no = 1),
        status = as.factor(status)
    )


```


```{r message = FALSE, warning=FALSE}
### Adding community district data
inc_df = read_csv("./Med_income_2017.csv") %>% 
    janitor::clean_names() %>% 
    mutate(
        pop_1000s = round(total_population/1000, 0),
        inc_1000s = round(median_income/1000, 1),
        income_bracket = case_when(
            median_income >= 20000 & median_income <= 30000 ~ "20-30k",
            median_income > 30000 & median_income <= 40000 ~ "30-40k",
            median_income > 40000 & median_income <= 50000 ~ "40-50k",
            median_income > 50000 & median_income <= 60000 ~ "50-60k",
            median_income > 60000 & median_income <= 70000 ~ "60-70k",
            median_income > 70000 & median_income <= 80000 ~ "70-80k",
            median_income > 80000 & median_income <= 90000 ~ "80-90k",
            median_income > 90000 & median_income <= 100000 ~ "90-100k",
            median_income > 100000 & median_income <= 125000 ~ "100-125k",
            median_income > 125000 & median_income <= 150000 ~ "125k+",
        ),
        income_bracket = as.factor(income_bracket)
    )

# adding income to data and removing any observations that do not have a specific community board to link to income
add_inc = left_join(nyc_tidy, inc_df, by = "community_board") %>% 
    filter(is.na(median_income) == FALSE) %>% 
    mutate(
        year_fac = as.factor(created_year)
    )


```



```{r}
# Calculate days to close
#data with days to close information
nyc_data <- add_inc %>% 
 filter(status == "Closed" & closed_year <= 2018) %>% 
    mutate(
        closed_year = as.numeric(closed_year),
        closed_month = as.numeric(closed_month),
        closed_day = as.numeric(closed_day)
    ) %>% 
    mutate(
        diff_year = closed_year - created_year,
        diff_month = closed_month - created_month,
        diff_day = closed_day - created_day,
        diff_dayt =  diff_year*365 + diff_month*30 + diff_day
       # diff_dayt = as.factor(diff_dayt)
  ) %>% 
  mutate(
    status = as.character(status)
  ) %>% 
  mutate(
    income_bracket = forcats::fct_relevel(income_bracket, "125k+", "100-125k", "90-100k", "80-90k", "70-80k", "60-70k", "50-60k", "40-50k", "30-40k", "20-30k")
  ) %>% 
  drop_na(closed_year)
  
```
# Tables

Tables report median, and IQR given the skewness of our data and the count. 
```{r, message=FALSE, warning=FALSE}
mycontrols  <- tableby.control(test=F, total=T,
                               cat.simplify = F,
                               numeric.stats=c("N", "median", "q1q3"),
                               cat.stats=c("countpct"),
                               stats.labels=list(N='Count', median='Median', q1q3='Q1,Q3'),
                               digits = 1)

tab_clo<- tableby(created_year ~ diff_dayt + status  + borough, data = nyc_data, control = mycontrols)

tab_inc<- tableby(created_year ~ income_bracket, data = nyc_data, control = mycontrols)

tab_comp <- tableby(created_year ~ complaint_simp, data = nyc_data, control = mycontrols)

tab_com_clo_inc <- tableby(complaint_simp ~ diff_dayt +income_bracket + borough , data = nyc_data, control = mycontrols)

```

In table 1 the median and interquartile range of how many days it took for cases to close.  Given the data is skewed right I reported the median values being `1` to `2` days and largest interqutile range being `0 - 9` and smallest being `0 - 6` acroos years `2014 - 2018`. 

Most complaints are closed within a week with a high density of them being closed within a day or two. Additionally, we observe even distribution across all boroughs and closed cases. 
```{r, results="asis" }
summary(tab_clo, title = "NYC 311 Complaints Demographics: table 1")

```
In this table we see that individuals in the income bracket 40 - 50k (25.4%) made the most complaints through out all the years in this study while individuals in the 90 - 100k (2.6%)made the least comlaints consistently. 
```{r, results="asis" }
summary(tab_inc, title = "NYC 311 Complaints Demographics: table 2 ")
```
In table 2 we see that the most complaints are of noise (17.9%).  Interestingly we see that car/traffic complaints increase overtime that suggest an increase in conjestion. 
```{r,results="asis" }
summary(tab_comp, title = "NYC 311 Complaints Demographics: table 3 ")

```

In this table  cases with complaints that have something to do with tree, paint/plaster, and maintanance take the longest to close 11 - 12 days compared to the other compalaint cases that close with the said interquartile range of 0-9 days.  Taking a look at the incomme bracket we observe that higher income brackest mostly complain about the homeless (17.2%) while the lowest income clase mostly complain about heat (20.3%) and maintanance(20.9%) comparing accross all income brackets. Furthermore we can also see that Manhattan has most of its complaints related to air quality (38.8%), hazardous material (30.3%), homeless (77.4%), and noise (30.5%) comparing across all boroughs.  

```{r,results="asis" }
summary(tab_com_clo_inc , title = "NYC 311 Complaints Demographics: table 4 ")

```


