---
title: "NYC Open Data"
author: "Steven Lawrence"
date: "November 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#library(RSocrata)
library(tidyverse)

```

This token is to beable to acces data from the socrata repository.  
```{r,eval=F}
token <- "QsuUJQpyjGciXV5KibPhY64hO"
```
Here is an example to pull data from the API.  the `url` includes the endpoint for the dataset ending with a `?`. Subsequently, using the variable names a query functions to organize the data ie: `$limit` and `borough=BROOKLYN`. However this did not work as I planned so I described how I collected the data below. 
```{r,eval=F}
brooklyn<-RSocrata::read.socrata(
    url="https://data.cityofnewyork.us/resource/erm2-nwe9.csv?borough=BROOKLYN&$created_date between `2014/04/23 12:00:00 and 2018/03/05 12:00:00`&$limit=100", 
    app_token = token) 

write_csv(brooklyn, "brooklyn_311_100k.csv")

brooklyn_csv <- read_csv("brooklyn_311_100k.csv")

dim(brooklyn)

head(brooklyn_csv)

```

#Importing and merging data

I downloaded the data from the NYC Open data website filtering for each boorough and time frame 2014 - 2018.  Then I sampled from each dataset, containing 700k - 3 million rows, 100k data points in r, cleaned the `created_date` variable, and checked for equal sampling across the years. 

Brooklyn

```{r,eval=F}

read_csv("311_brooklyn.csv") %>%
    sample_n(100000, replace = F) %>%
     separate(`Created Date`, into = c("month","day","year"), sep = "\\/" ) %>%
    separate(year, into = c("year","time"), sep = " ") %>%
    write_csv(., "brooklyn_311_100k.csv")

brooklyn<- read_csv("brooklyn_311_100k.csv")

table(brooklyn100k$year)

```

Manhattan
```{r,eval=F}

read_csv("311_manhattan.csv") %>%
    sample_n(100000, replace = F) %>%
     separate(`Created Date`, into = c("month","day","year"), sep = "\\/" ) %>%
    separate(year, into = c("year","time"), sep = " ") %>%
    write_csv(., "manhattan_311_100k.csv")

manhattan<- read_csv("manhattan_311_100k.csv")

table(manhattan$year)

```

Queens
```{r,eval=F}

read_csv("311_queens.csv")%>%
    sample_n(100000, replace = F) %>%
     separate(`Created Date`, into = c("month","day","year"), sep = "\\/" ) %>%
    separate(year, into = c("year","time"), sep = " ") %>%
    write_csv(., "queens_311_100k")

queens<- read_csv("queens_311_100k")

table(queens$year)

```
Bronx
```{r,eval=F}

read_csv("311_bronx.csv")%>%
    sample_n(100000, replace = F) %>%
     separate(`Created Date`, into = c("month","day","year"), sep = "\\/" ) %>%
    separate(year, into = c("year","time"), sep = " ") %>%
    write_csv(., "bronx_311_100k.csv")

bronx<- read_csv("bronx_311_100k.csv")

table(bronx$year)



```
Staten Island

```{r,eval=F}
 

read_csv("311_statenisland.csv")  %>%
    sample_n(100000, replace = F) %>%
     separate(`Created Date`, into = c("month","day","year"), sep = "\\/" ) %>%
    separate(year, into = c("year","time"), sep = " ") %>%
    write_csv(., "statenisland_311_100k.csv")

statenisland <- read_csv("statenisland_311_100k.csv")

table(statenisland$year)


```
Unspecified
```{r,eval=F}


read_csv("311_unspecified.csv")  %>%
    sample_n(100000, replace = F) %>%
     separate(`Created Date`, into = c("month","day","year"), sep = "\\/" ) %>%
    separate(year, into = c("year","time"), sep = " ") %>%
    write_csv(.,"unspecified_311_100k")

unspec <- read_csv("unspecified_311_100k")

head(unspec)

table(unspec$year)


```

Combinging all the datasets of 100k data points.
```{r, eval=F}
rbind(brooklyn, manhattan, queens, bronx, statenisland, unspec) %>%
    rename(created_year = "year" , created_day = "day",created_month = "month",created_time = "time") %>%
    janitor::clean_names() %>%
   # write_csv(., "p8105nyc_311_100k.csv")
```

```{r}
nyc_311<-read_csv("p8105nyc_311_100k.csv") 

nyc_311 %>% 

```

```{r}
nyc_311 %>% 
    
```

