---
title: "regression_analysis"
author: "Ava Hamilton"
date: "12/1/2019"
output: html_document
---

```{r setup, include=FALSE}
#devtools::install_github("Chicago/RSocrata")
library(RSocrata)
library(tidyverse)

```

## read in data
```{r message = FALSE}
nyc <- read_csv(file = "../p8105nyc_311_100k.csv") %>% 
    janitor::clean_names()

nyc_tidy <- nyc %>%   
    filter(borough != "Unspecified") %>% 
    separate(closed_date, 
             into = c("closed_month","closed_day","closed_year"), 
             sep = "\\/" ) %>%
    separate(closed_year, 
             into = c("closed_year","closed_time"), 
             sep = " ") %>% 
    mutate(
        created_year = as.numeric(created_year),
        created_month = as.numeric(created_month),
        created_day = as.numeric(created_day),
        city = as.factor(city),
        status = as.factor(status),
        borough = as.factor(borough),
        agency = as.factor(agency),
        complaint_type = as.factor(complaint_type),
        community_board = as.factor(community_board),
        open_complaint = ifelse(status == "Closed", yes = 0, no = 1),
        # open_complaint = ifelse(is.na(closed_year),  yes = 1, no = 0),
        complaint_simp = case_when(
            str_detect(complaint_type, 
                       regex("street", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("sidewalk", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("curb", ignore_case = TRUE)) ~ "Street Condition",
            str_detect(complaint_type, 
                       regex("noise", ignore_case = TRUE)) ~ "Noise",
            str_detect(complaint_type, 
                       regex("heat", ignore_case = TRUE)) ~ "Heat",
            str_detect(complaint_type, 
                       regex("water", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("leak", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("plumbing", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("boiler", ignore_case = TRUE)) ~ "Water/plumbing",
            str_detect(complaint_type, 
                       regex("paint", ignore_case = TRUE)) ~ "Paint/Plaster",
            str_detect(complaint_type, 
                       regex("asbestos", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("lead", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("hazard", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("mold", ignore_case = TRUE)) ~ "Hazard Material",
            str_detect(complaint_type, 
                       regex("elevator", ignore_case = TRUE)) 
            |str_detect(complaint_type, 
                        regex("maintenance", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("electric", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("stairs", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("window", ignore_case = TRUE)) 
            |str_detect(complaint_type, 
                        regex("appliance", ignore_case = TRUE)) ~ "Maintenance",
            str_detect(complaint_type, 
                       regex("sanita", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("rodent", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("dirty", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("sew", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("standing water", ignore_case = TRUE)) ~ "Sanitation",
            str_detect(complaint_type, 
                       regex("tree", ignore_case = TRUE)) ~ "Tree",
            str_detect(complaint_type, 
                       regex("parking", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("car", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("drive", ignore_case = TRUE))
            |str_detect(complaint_type,  
                        regex("vehicle", ignore_case = TRUE))
            |str_detect(complaint_type,  
                        regex("traffic", ignore_case = TRUE)) ~ "Car/Traffic",
            str_detect(complaint_type, 
                       regex("air", ignore_case = TRUE)) ~ "Air Quality",
            str_detect(complaint_type, 
                       regex("collection", ignore_case = TRUE)) ~ "Collection",
            str_detect(complaint_type, 
                       regex("homeless", ignore_case = TRUE))
            |str_detect(complaint_type, 
                        regex("panhandling", ignore_case = TRUE)) ~ "Homeless"),
        health_complaint = ifelse(
            complaint_simp %in% c("Heat", "Water/Plumbing", "Hazard Material", "Sanitation", "Air Quality"), yes = 1, no = 0),
        complaint_simp = as.factor(complaint_simp),
        open_health_complaint = case_when(
            open_complaint == 1 & health_complaint == 1 ~ 1,
            open_complaint == 0 | health_complaint == 0 ~ 0
        ),
        # openCorr = ifelse(status == "Closed", yes = 0, no = 1),
        status = as.factor(status)
    )


```

Adding community district data
```{r message = FALSE}

inc_df = read_csv("./Med_income_2017.csv") %>% 
    janitor::clean_names() %>% 
    mutate(
        pop_1000s = round(total_population/1000, 0),
        inc_1000s = round(median_income/1000, 1),
        income_bracket = case_when(
            median_income >= 20000 & median_income <= 30000 ~ "20k-30k",
            median_income > 30000 & median_income <= 40000 ~ "30-40k",
            median_income > 40000 & median_income <= 50000 ~ "40-50k",
            median_income > 50000 & median_income <= 60000 ~ "50-60k",
            median_income > 60000 & median_income <= 70000 ~ "60-70k",
            median_income > 70000 & median_income <= 80000 ~ "70-80k",
            median_income > 80000 & median_income <= 90000 ~ "80-90k",
            median_income > 90000 & median_income <= 100000 ~ "90-100k",
            median_income > 100000 & median_income <= 125000 ~ "100-125k",
            median_income > 125000 & median_income <= 150000 ~ "125k+",
        ),
        income_bracket = as.factor(income_bracket)
    )

# adding income to data and removing any observations that do not have a specific community board to link to income
add_inc = left_join(nyc_tidy, inc_df, by = "community_board") %>% 
    filter(is.na(median_income) == FALSE) %>% 
    mutate(
        year_fac = as.factor(created_year)
    )


```

### Aggregating data to community district and year level

To examine the effect of community district level variables on total number of complaints we must first aggregate to the community district level. We are also interested in how year effects the number of complaints in each district so we group by year as well to produce complaint totals for each year and community district.

Once we have complaint totals, we do not want individual complaint information anymore. We select distinct year * community district observations. 
```{r message = FALSE}

# grouping by community district and year
cb_group_year = add_inc %>% 
    group_by(community_board, created_year) %>%
    add_count(community_board, name = "number_complaints") %>% 
    mutate(
        num_unsolved = sum(open_complaint),
        num_health_complaint = sum(health_complaint),
        num_open_health = sum(open_health_complaint)
    ) %>% 
    select(number_complaints, num_unsolved, num_open_health, everything())


########## including year
cb_group_year_distinct = cb_group_year %>% 
    select(community_board, number_complaints, inc_1000s, num_unsolved, num_health_complaint, borough, per_black_nh, per_hisp, per_white_nh, median_income, num_open_health, created_year, year_fac, pop_1000s) %>% 
    distinct() %>% 
    arrange(community_board) %>% 
    mutate(
        borough = relevel(borough, ref = "MANHATTAN")
        )

cb_group_year_distinct = within(cb_group_year_distinct, borough <- relevel(borough, ref = "MANHATTAN"))

manhanttan_distinct = cb_group_year_distinct %>% 
    filter(borough == "MANHATTAN")

brooklyn_distinct = cb_group_year_distinct %>% 
    filter(borough == "BROOKLYN")

bronx_distinct = cb_group_year_distinct %>% 
    filter(borough == "BRONX")

queens_distinct = cb_group_year_distinct %>% 
    filter(borough == "QUEENS")

SI_distinct = cb_group_year_distinct %>% 
    filter(borough == "STATEN ISLAND")


```

## What neighborhood characteristics effect number and type of complaints in each community district? 

We conducted linear models to examine predictors of number of complaints from 2014-2018 per community district. Our three outcomes were:

1) Total number of complaints
2) Number of health complaints
3) Number of unresolved complaints.

We looked at these categories of complaints from 2014-2018 predicted by year, and community district level variables including median income (in the 1000s), total population (in the 1000s), percent non-hispanic black and percent Hispanic. For outcomes 1 and 2 we included number of unresolved complaints as a predictor as well. 

As we have equal complaints across borough we startified our analysis to predict number of complaints within each borough.

#### 1) Number of complaints by NYC community district:

##### **A) Bronx**

```{r message = FALSE}

# number of complaints
options(scipen = 5)

nom_comp_bronx = lm(number_complaints ~ inc_1000s + num_unsolved + per_black_nh + per_hisp + pop_1000s +created_year, data = bronx_distinct)
summary(nom_comp_bronx)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_bronx) %>% broom::glance() %>% knitr::kable()

```

##### **B) Brooklyn**

```{r message = FALSE}
# Brooklyn
nom_comp_brook = lm(number_complaints ~ inc_1000s + num_unsolved + per_black_nh + per_hisp + pop_1000s +created_year, data = brooklyn_distinct)
summary(nom_comp_brook)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_brook) %>% broom::glance() %>% knitr::kable()

```

##### **C) Manhattan**

```{r message = FALSE}
# Manhattan

nom_comp_manhattan = lm(number_complaints ~ inc_1000s + num_unsolved + per_black_nh + per_hisp + pop_1000s +created_year, data = manhanttan_distinct)
summary(nom_comp_manhattan)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_manhattan) %>% broom::glance() %>% knitr::kable()

```

##### **D) Queens**

```{r message = FALSE}
# Queens
nom_comp_queens = lm(number_complaints ~ inc_1000s + num_unsolved + per_black_nh + per_hisp + pop_1000s +created_year, data = queens_distinct)
summary(nom_comp_queens)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_queens) %>% broom::glance() %>% knitr::kable()

```

##### **E) Staten Island**

```{r message = FALSE}
# SI
nom_comp_SI = lm(number_complaints ~ inc_1000s + num_unsolved + per_black_nh +created_year, data = SI_distinct)
summary(nom_comp_SI)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_SI) %>% broom::glance() %>% knitr::kable()

```

##### **Residuals**


```{r}

par(mfrow = c(2,3))
plot(nom_comp_bronx, which = 1, main = "Bronx
")
plot(nom_comp_brook, which = 1, main = "Brooklyn
")
plot(nom_comp_manhattan, which = 1, main = "Manhattan
")
plot(nom_comp_queens, which = 1, main = "Queens
")
plot(nom_comp_SI, which = 1, main = "Staten Island
")


```



*Comments:*

In the Bronx, on average, for each additional year we see `r round(pull((summary(nom_comp_bronx)%>% broom::tidy()) %>% filter(term == "created_year"),estimate),0)` more complaints in each community district, for each additional 1000 people in a CD we see `r round(pull((summary(nom_comp_bronx)  %>% broom::tidy()) %>% filter(term == "pop_1000s"),estimate),0)` more complaints, and for each $1000 increase in median income we see `r -1*round(pull((summary(nom_comp_bronx)  %>% broom::tidy()) %>% filter(term == "inc_1000s"),estimate),0)` *_fewer_* complaints.

In Manhattan and Queens we see an effect on total number of complaints by number of unresolved complaints. On average, for each additional unresolved complaint people in Manhattan and Queens will have `r round(pull((summary(nom_comp_manhattan)  %>% broom::tidy()) %>% filter(term == "num_unsolved"),estimate),0)` and `r round(pull((summary(nom_comp_queens) %>% broom::tidy()) %>% filter(term == "num_unsolved"),estimate),0)` more complaints respectively.

Based on the adjusted R-squared values, number of complaints in a CD range from  `r 100*round(pull((summary(nom_comp_manhattan) %>% broom::glance()),adj.r.squared),4)`% in Manhattan to `r 100*round(pull((summary(nom_comp_SI) %>% broom::glance()),adj.r.squared),4)`% in Staten Island predicted by this combination of variables. Residuals appear to follow normal distributions.



#### 2) Number of health-related complaints by NYC community district:

##### **A) Bronx**

```{r message = FALSE}

# number of complaints
options(scipen = 5)

nom_comp_h_bronx = lm(num_health_complaint ~  inc_1000s + num_unsolved + per_black_nh + per_hisp + pop_1000s + created_year, data = bronx_distinct)
summary(nom_comp_h_bronx)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_h_bronx) %>% broom::glance() %>% knitr::kable()

```

##### **B) Brooklyn**

```{r message = FALSE}
# Brooklyn
nom_comp_h_brook = lm(num_health_complaint ~ inc_1000s + num_unsolved + per_black_nh + per_hisp + pop_1000s +created_year, data = brooklyn_distinct)
summary(nom_comp_h_brook)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_h_brook) %>% broom::glance() %>% knitr::kable()

```

##### **C) Manhattan**

```{r message = FALSE}
# Manhattan

nom_comp_h_manhattan = lm(num_health_complaint ~ inc_1000s + num_unsolved + per_black_nh + per_hisp + pop_1000s +created_year, data = manhanttan_distinct)
summary(nom_comp_h_manhattan)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_h_manhattan) %>% broom::glance() %>% knitr::kable()

```

##### **D) Queens**

```{r message = FALSE}
# Queens
nom_comp_h_queens = lm(num_health_complaint ~ inc_1000s + num_unsolved + per_black_nh + per_hisp + pop_1000s +created_year, data = queens_distinct)
summary(nom_comp_h_queens)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_h_queens) %>% broom::glance() %>% knitr::kable()

```

##### **E) Staten Island**

```{r message = FALSE}
# SI
nom_comp_h_SI = lm(num_health_complaint ~ inc_1000s + num_unsolved + per_black_nh +created_year, data = SI_distinct)
summary(nom_comp_h_SI)  %>% broom::tidy() %>% knitr::kable()
summary(nom_comp_h_SI) %>% broom::glance() %>% knitr::kable()

```

##### **Residuals**


```{r}

par(mfrow = c(2,3))
plot(nom_comp_h_bronx, which = 1, main = "Bronx
")
plot(nom_comp_h_brook, which = 1, main = "Brooklyn
")
plot(nom_comp_h_manhattan, which = 1, main = "Manhattan
")
plot(nom_comp_h_queens, which = 1, main = "Queens
")
plot(nom_comp_h_SI, which = 1, main = "Staten Island
")


```



*Comments:*

Complaints were determined to be health related if they were about heat, hot water, sanitary conditions, poor air quality, or hazardous materials.

In Manhattan and Queens we see a significant effect on total number of health related complaints by median income of the CD. On average, for each $1000 increase in median income we see `r -1*round(pull((summary(nom_comp_h_manhattan)  %>% broom::tidy()) %>% filter(term == "inc_1000s"),estimate),0)` more health-related complaints in Manhattan `r -1*round(pull((summary(nom_comp_h_queens)  %>% broom::tidy()) %>% filter(term == "inc_1000s"),estimate),0)` *_fewer_* health-related complaints in Queens.

Based on the adjusted R-squared values, number of health related complaints in a CD range from  `r 100*round(pull((summary(nom_comp_h_brook) %>% broom::glance()),adj.r.squared),4)`% in Brooklyn to `r 100*round(pull((summary(nom_comp_h_SI) %>% broom::glance()),adj.r.squared),4)`% in Staten Island predicted by this combination of variables. The plots of the residuals indicate that future analyses should increase or adjust the predictors to better fit the data.


#### 3) Number of unresolved complaints by NYC community district:

##### **A) Bronx**

```{r message = FALSE}

# number of complaints
options(scipen = 5)

nom_unres_bronx = lm(num_unsolved ~ inc_1000s + per_black_nh + per_hisp + pop_1000s +created_year, data = bronx_distinct)
summary(nom_unres_bronx)  %>% broom::tidy() %>% knitr::kable()
summary(nom_unres_bronx) %>% broom::glance() %>% knitr::kable()

```

##### **B) Brooklyn**

```{r message = FALSE}
# Brooklyn
nom_unres_brook = lm(num_unsolved ~ inc_1000s + per_black_nh + per_hisp + pop_1000s +created_year, data = brooklyn_distinct)
summary(nom_unres_brook)  %>% broom::tidy() %>% knitr::kable()
summary(nom_unres_brook) %>% broom::glance() %>% knitr::kable()

```

##### **C) Manhattan**

```{r message = FALSE}
# Manhattan

nom_unres_manhattan = lm(num_unsolved ~ inc_1000s + per_black_nh + per_hisp + pop_1000s +created_year, data = manhanttan_distinct)
summary(nom_unres_manhattan)  %>% broom::tidy() %>% knitr::kable()
summary(nom_unres_manhattan) %>% broom::glance() %>% knitr::kable()

```

##### **D) Queens**

```{r message = FALSE}
# Queens
nom_unres_queens = lm(num_unsolved ~ inc_1000s + per_black_nh + per_hisp + pop_1000s +created_year, data = queens_distinct)
summary(nom_unres_queens)  %>% broom::tidy() %>% knitr::kable()
summary(nom_unres_queens) %>% broom::glance() %>% knitr::kable()

```

##### **E) Staten Island**

```{r message = FALSE}
# SI
nom_unres_SI = lm(num_unsolved ~ inc_1000s + per_black_nh +created_year, data = SI_distinct)
summary(nom_unres_SI)  %>% broom::tidy() %>% knitr::kable()
summary(nom_unres_SI) %>% broom::glance() %>% knitr::kable()

```

##### **Residuals**


```{r}

par(mfrow = c(2,3))
plot(nom_unres_bronx, which = 1, main = "Bronx
")
plot(nom_unres_brook, which = 1, main = "Brooklyn
")
plot(nom_unres_manhattan, which = 1, main = "Manhattan
")
plot(nom_unres_queens, which = 1, main = "Queens
")
plot(nom_unres_SI, which = 1, main = "Staten Island
")


```



*Comments:*


A complaint was labeled unresolved if its status was not labeled as "closed". For all boroughs but Staten Island, year is, on average, a significant predictor if number of unresolved complaints for each CD. 

In Brooklyn and Queens we see significant effect on total number of unresolved complaints by the percent non-Hispanic Black population. On average, we see `r 5*round(pull((summary(nom_unres_brook)  %>% broom::tidy()) %>% filter(term == "per_black_nh"),estimate),0)` and `r 5*round(pull((summary(nom_unres_queens)  %>% broom::tidy()) %>% filter(term == "per_black_nh"),estimate),0)` more unresolved complaints for each additonal 5% increase in non-Hispanic Black population in Brooklyn and Queens respectively.

Based on the adjusted R-squared values, number of unresolved complaints in a CD range from  `r 100*round(pull((summary(nom_unres_SI) %>% broom::glance()),adj.r.squared),4)`% in Staten Island to `r 100*round(pull((summary(nom_unres_brook) %>% broom::glance()),adj.r.squared),4)`% in The Bronx predicted by this combination of variables. The plots of the residuals indicate that future analyses should increase or adjust the predictors to better fit the data.


## Discussion

Between 2014 and 2018, approximately 22 million 311 complaints were made across the 5 boroughs in NYC. We were interested in how complaints differ by borough and community district. Based on our descriptive figures, mapped data, and linear models of NYC 311 complaint data, we see that there are a few key factors that play into how types of complaints are dispersed across the city.

With approximately 5000 311 complaints made everyday, a majority of which close within the week (66%), it is interesting to look into where the open complaints remain.